When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-06-25 doc/paper/abstract.tex
refactored to have independent, single, abstract file.

\documentclass[12pt]{article}
\input{def.tex}

\begin{document}

\title{When does vapor pressure deficit drive or reduce
  evapotranspiration?}

\author[1]{Adam Massmann\thanks{Corresponding author: akm2203@columbia.edu}}
\author[1]{Pierre Gentine}
\author[1,2]{Changjie Lin}

\affil[1]{Department of Earth and Environmental Engineering, Columbia University, New York, NY 10027}
\affil[2]{State Key Laboratory of Hydroscience and Engineering, Department of Hydraulic
  Engineering, Tsinghua University, Beijing, CN 100084}

\maketitle

\begin{abstract}
  Increasing vapor pressure deficit increases atmospheric demand for water, and vapor pressure deficit is expected to rise with increasing greenhouse gases. While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effort to conserve water. Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET. We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.

 The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD. Plants that are evolved or bred to prioritize primary production over water conservation (e.g. crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing). However for forest, grass, savannah, and shrub plant types, ET more frequently decreases than increases with rising VPD. This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world. 
\end{abstract}

\onehalfspacing

\input{body.tex}

>>>>

Increasing vapor pressure deficit (VPD) increases atmospheric demand for
water, and vapor pressure deficit is expected to rise with increasing
greenhouse gases. While increased evapotranspiration (ET) in response to
increased atmospheric demand seems intuitive, plants are capable of
reducing ET in response to increased VPD by closing their stomata, in an
effort to conserve water. Here we examine which effect dominates
response to increasing VPD: atmospheric demand and increases in ET, or
plant physiological response (stomata closure) and decreases in ET. We
use Penman-Monteith, combined with semi-empirical optimal stomatal
regulation theory and underlying water use efficiency, to develop a
theoretical framework for understanding how ET responds to increases in
VPD.

The theory suggests that for most environmental conditions and plant
types, plant physiological response dominates and ET decreases with
increasing VPD. Plants that are evolved or bred to prioritize primary
production over water conservation (e.g. crops) exhibit a higher
likelihood of atmospheric demand-driven response (ET
increasing). However for forest, grass, savannah, and shrub plant types,
ET more frequently decreases than increases with rising VPD. This work
serves as an example of the utility of our simplified framework for
disentangling land-atmosphere feedbacks at multiple scales, including
the characterization of ET response in an atmospherically drier,
enriched CO$_2$ world.

### 2018-06-25 doc/paper/vpd_et_paper_afm.tex
figs, data and citations now match data policy, just need to proofread body and then can submit v2 to arxiv
refactored to have independent, single, abstract file.

% 11/23/2015
\documentclass[draft,linenumbers]{afmjournal}
% \draftfalse
\usepackage{makecell} \usepackage{multirow} %AM for tables
\usepackage{booktabs} \drafttrue

@@

\begin{abstract}
  Increasing vapor pressure deficit increases atmospheric demand for  water, and vapor pressure deficit is expected to rise with  increasing greenhouse gases. While increased evapotranspiration (ET) in  response to increased atmospheric demand seems intuitive, plants are  capable of reducing ET in response to increased VPD by closing their  stomata, in an effort to conserve water. Here we examine which  effect dominates response to increasing VPD: atmospheric demand and  increases in ET, or plant physiological response (stomata closure)  and decreases in ET. We use Penman-Monteith, combined with  semi-empirical optimal stomatal regulation theory and underlying  water use efficiency, to develop a theoretical framework for  understanding how ET responds to increases in VPD.

  The theory suggests that for most environmental conditions and plant  types, plant physiological response dominates and ET decreases with  increasing VPD. Plants that are evolved or bred to prioritize  primary production over water conservation (e.g. crops) exhibit a  higher likelihood of atmospheric demand-driven response (ET  increasing). However for forest, grass, savannah, and shrub plant  types, ET more frequently decreases than increases with rising  VPD. This work serves as an example of the utility of  our simplified framework for disentangling land-atmosphere feedbacks  at multiple scales, including the characterization of ET response in  an atmospherically drier, enriched CO$_2$ world. 
\end{abstract}

>>>>

% 11/23/2015
\documentclass[draft,linenumbers]{afmjournal}
% \draftfalse
\usepackage{makecell} \usepackage{multirow} %AM for tables
\usepackage{hyperref}
\usepackage[title]{appendix}
\usepackage{booktabs} \drafttrue
\usepackage{longtable}

@@

\begin{abstract}
\input{abstract.tex}
\end{abstract}
